A comparison of classification algorithms applied to the Adult Database

### 2009-12-10 report/final.tex
more bayesian networks

The experiments conducted will use genetic based search, hill climbing \cite{bayes3}, simulated annealing,
and a modified hill climbing algorithm, named K2 \cite{bayes2}.

@@

This section shows performance results for the various algorithms.

@@

From the table we conclude that we need a certain amount of trees ($\approx$80) to
reach a satisfactory error rate ($<$22\%) for the test set. Interestingly
the error rate for the train set is low ($\approx$9\%), which indicates over-fitting.

One drawback of using a big number of trees is the use of large
quantities of memory (we needed 1GB to compute with 400 trees) and the time needed to compute everything.

@@

The CN2 software used supports different algorithms for building either ordered
or unordered rule sets and with different error estimations: laplacian, naive,
information gain and modified entropy. We've tested each combination of
parameters and the overall results are presented in table \ref{tbl:results_cn2}.

\begin{table}
\begin{center}
\begin{tabular}{ | l | l | l | l | l | }
    \hline
    \textbf{Algorithm} & \textbf{Error Estimate} & \textbf{Train Error Rate} & \textbf{Test Error Rate} & \textbf{Time}
\end{tabular}
\caption{Results for CN2.}
\label{tbl:results_cn2}
\end{center}
\end{table}

@@

\bibitem{bayes2}
G. Cooper, E. Herskovits. A Bayesian method for the induction of probabilistic networks from 
data. Machine Learning, 9: 309–347, 1992.

>>>>

The experiments conducted will use genetic based search, hill climbing \cite{bayes3}, and simulated annealing.

@@

This section shows performance results for the various algorithms described in section \ref{sec:algs}.

@@

From the table we conclude that we need a certain amount of trees ($\approx$80) to
reach a satisfactory error rate ($<$22\%) for the test set. Interestingly
the error rate for the train set is low ($\approx$9\%), which indicates over-fitting
when comparing the difference to the test error rate.

One drawback of using a big number of trees is the use of large
quantities of memory (we needed 1GB to compute 400 trees) and the long time needed to construct the classifier.

@@

\subsection{Bayesian Networks}

Our strategy for bayesian networks revolve around changing the search algorithm
and its parameters. Weka easily allows the parameterization of this algorithm (\textbf{$<$-Q$>$})
that optimizes the network structure.

The search algorithms experimented include:

\begin{itemize}
  \item Genetic based search (\textit{weka.classifiers.bayes.net.search.local.GeneticSearch}) \\
  The following options are available:
  \begin{itemize}
    \item -L integer: Population size. 
    \item -A integer: Descendant population size.
    \item -U integer: Number of runs/generations.
    \item -M: Use mutation (true by default).
    \item -C: Use cross-over (true by default).
  \end{itemize}

  \item Hill climbing (\textit{weka.classifiers.bayes.net.search.local.HillClimber}) \\
  The following options are available:
  \begin{itemize}
    \item -P $<$nr of parents$>$: Maximum number of parents.
    \item -R: Use arc reversal operation. (default false)
  \end{itemize}

  \item Simulated annealing (\textit{weka.classifiers.bayes.net.search.local.SimulatedAnnealing}) \\
  The following options are available:
  \begin{itemize}
    \item -A float: Start temperature.
    \item -U integer: Number of runs.
    \item -D float: Delta temperature.
  \end{itemize}
\end{itemize}

Weka also allows the parameterization of the estimator algorithm (\textbf{$<$-E$>$}).
We will use the \textit{SimpleEstimator} estimator (\textit{weka.classifiers.bayes.net.estimate.SimpleEstimator -- -A 1.0})
for all the experiments.

\subsubsection{Genetic Search}

The Genetic Search algorithm is a genetic based algorithm to optimize the bayesian network structure.
We can configure the population size and number of generations. Mutation and cross-over can also
be set or kept out.

For this experiment we will only use a portion of the train data to speed up the process.

\begin{table}[ht]
  \begin{center}
  \begin{tabular}{ | l | c | c | c | c | c | c |}
    \hline
    \textbf{Parameter} & \textbf{Portion} & \textbf{Portion error rate} & \textbf{Train error rate} & \textbf{Test error rate} & \textbf{Time} \\ \hline
    GeneticSearch -C -M -L 3 -A 3 -U 3 & 5 & & & & \\ \hline
    GeneticSearch -C -M -L 3 -A 3 -U 5 & 5 & & & & \\ \hline
    GeneticSearch -C -M -L 3 -A 3 -U 15 & 5 & & & & \\ \hline
  \end{tabular}
  \caption{Results for Bayesian Networks / Genetic Search.}
  \label{tbl:results_bayesian_networks_gs}
  \end{center}
\end{table}

\subsubsection{Hill Climbing}

The best configuration for hill climbing (Table \ref{tbl:results_bayesian_networks_hc})
is when the number of parents is exactly 1, still, the error rate is somewhat unsatisfactory when
comparing to other results.

Incrementing the number of parents forces the classifier to over-fit the train set, as it
directly increases the test set error rate while decreasing the train set error rate.

\begin{table}[ht]
  \begin{center}
  \begin{tabular}{ | l | c | c | c | c |}
    \hline
    \textbf{Parameters} & \textbf{Train error rate} & \textbf{Test error rate} & \textbf{Time} \\ \hline
    HillClimber -P 1 & 19.22 & 22.93 & 0m29.512s \\ \hline
    HillClimber -P 2 & 15.15 & 26.34 & 0m30.340s \\ \hline
    HillClimber -P 3 & 14.51 & 28.11 & 0m29.963s \\ \hline
    HillClimber -P 4 & 14.43 & 27.92 & 0m29.945s \\ \hline
    HillClimber -P 10 & 14.43 & 27.92 & 0m30.638s \\ \hline
  \end{tabular}
  \caption{Results for Bayesian Networks / Hill Climbing.}
  \label{tbl:results_bayesian_networks_hc}
  \end{center}
\end{table}

\subsubsection{Simulated Annealing}

The simulated annealing algorithm provides lots of configuration parameters
and it is difficult to input the best values of start temperature, delta
temperature and number of runs to archive the best error rates.

We tried various temperatures values and came up with the Table \ref{tbl:results_bayesian_networks_sa}.

\begin{table}[ht]
  \begin{center}
  \begin{tabular}{ | l | c | c | c | c |}
    \hline
    \textbf{Parameters} & \textbf{Train error rate} & \textbf{Test error rate} & \textbf{Time} \\ \hline
    SimulatedAnnealing -U 75 -D 0.4 -A 20 & 14.32 & 26.80 & 0m29.841s \\ \hline
    SimulatedAnnealing -U 55 -D 0.4 -A 20 & 14.40 & 21.45 & 0m29.77s \\ \hline
    SimulatedAnnealing -U 52 -D 0.4 -A 20 & 14.49 & 21.09 & 0m29.407s \\ \hline
    SimulatedAnnealing -U 50 -D 0.4 -A 20 & 15.2 & 21.35 & 0m30.884s \\ \hline
    SimulatedAnnealing -U 50 -D 0.4 -A 40 & 15.2 & 21.35 & 0m30.884s \\ \hline
    SimulatedAnnealing -U 25 -D 0.4 -A 20 & 18.25 & 22.31 & 0m29.602s \\ \hline
    SimulatedAnnealing -U 10 -D 0.4 -A 20 & 19.82 & 22.96 & 0m29.131s \\ \hline
  \end{tabular}
  \caption{Results for Bayesian Networks / Simulated Annealing.}
  \label{tbl:results_bayesian_networks_sa}
  \end{center}
\end{table}

The diminish the error rates more research will be required to find the best
tuple of configuration values. Anyway, with just a few guesses we came up
with $\approx21$\% error rate, which is quite good when comparing to other algorithms.

@@
